An Early Eocene temperature gradient modelled with geological and ecological proxies

### 2022-10-07 manuscript/manuscript.Rmd
manuscript

Geochemical climate proxy data were extracted from a latest Paleocene and early Eocene compilation [@hollis2019]. This compilation provides  For our analyses, this dataset was restricted to the EECO and contains four commonly used geochemical climate proxies: d180, d47, Mg/Ca and TEX86 (REFS). For a detailed description of each see [@hollis2019].

@@

 Mangrove localities containing *Avicennia* pollen, but no Rhizophoraceae pollen, a mean temperature of 19.05 degC was ascribed. The absence of Rhizophoraceae indicates that sea surface temperature values

This value is between *Avicennia*'s lower temperature limit (15.6 degC) and the temperature at which *Rhizophora* commonly occurs (22.5 degC), deliberately wider than indicated by the lower *Rhizophora* limit.

To translate mangrove occurrences from the compilation by @Popescu2021 to a quantitative temperature proxy, we allocate locations at which *Avicennia* pollen, but no Rhizophoraceae pollen were found, a mean temperature of 19.05 degC, between the lower *Avicennia* limit (15.6 degC) and the temperature at which *Rhizophora* commonly occurs (22.5 degC), deliberately wider than indicated by the lower *Rhizophora* limit. We set the standard deviation to 1.725 such that 95 % of the probability density are within 15.6 and 22.5 degC, placing 5% probability on the possibility that an Eocene *Avicennia*-only mangrove record occurred outside of this range. We define a 95% probability density for locations with *Avicennia* and Rhizophoraceae in a similar way. To allow for greater uncertainty at the lower temperature limit, we use the empirical 20.7 degC limit. For the upper limit, we use the maximal coral reef temperature of 29.5 degC from @kleypas1999, as the compendium of @Quisthoudt2012 does not include the warmest, equatorial regions. The mean and standard deviation of locations with *Avicennia* and Rhizophoraceae are thus set to 25.1 and 2.2, respectively.

@@

The palaeogeographic distribution of geochemical and ecological data was reconstructed using the @merdith2021 plate rotation model via the palaeoverse R package ver. XXX (REF). The midpoint age of the EECO (51.5 Ma), along with the present-day coordinates of geochemical and ecological data, were used for palaeorotation.

@@

## EECO reconstruction
- Show Eocene gradient
- Global t
- report lat gradient

>>>>

Geochemical climate proxy data were extracted from a latest Paleocene and early Eocene compilation [@hollis2019]. This compilation provides  For our analyses, this dataset was restricted to the EECO and contains four commonly used geochemical climate proxies: d180, d47, Mg/Ca and TEX86 (REFS). For a detailed description of each proxy see [@hollis2019].

@@

@@

The palaeogeographic distribution of geochemical and ecological data was reconstructed using the @merdith2021 plate rotation model via the palaeoverse R package ver. XXX (REF). The midpoint age of the EECO (51.2 Ma), along with the present-day coordinates of geochemical and ecological data, were used for palaeorotation.

@@

## EECO reconstruction
- Show Eocene gradient
- Global t : use raster package for calculation of areal weights
- report lat gradient
